### 2016-04-29 itraquant/Benchmark/itraq-benchmark.Rmd
Bechmark

---
title: | 
  | Portable Proteomics Pipeline (P3) 
  | Labelled (iTRAQ4) Benchmark
author: "Daniel Kristiyanto (daniel.kristiyanto@pnnl.gov)"
date: \today
fontsize: 11pt
output: pdf_document
bibliography: ../../media/ref.bib
nocite: |
  @domon2006mass, @deutsch2010mass, @MSnba83
csl: ../../media/biomed-central.csl
header-includes:
  \usepackage{fancyhdr}
  \usepackage{graphicx}
  \usepackage{hyperref}
  \pagestyle{fancy}
  \setlength\headheight{28pt}
  \rhead{\includegraphics[width = .2\textwidth]{../../media/pnnllogo.png}}
  \fancyfoot[LE,RO]{kristiyanto/p3:benchmark}
---
\rule{\textwidth}{1pt}

```{r, warning=FALSE, echo=FALSE, include=FALSE}
#setwd("~/Documents/GITHUB/P3/itraquant/Benchmark/itraq-benchmark/")
load("~/Documents/GITHUB/P3/itraquant/Benchmark/itraq-benchmark/itraq-benchmark.RData")
library(MSnbase)
library(mzID)
cols. <- c()
cols <- c("accession", "pepseq", "ms-gf:evalue")

```

This document is to benchmark and as a sanity check to the iTRAQ4 method for the Portable Proteomics Pipelines (P3) containers.

## Results from PNNL
For this first purpose, `r length(mzid.files)` first file TCGA calculated by PNNL used as the comparasion:
```{r}
show(mzml.sp)
dim(msexp.sp) # The size of table
head(msexp.sp)[1:3,]
```

## Result from P3 container
For comparation, `r length(mzml.files )` file were also loaded and then merged with the identification data using the MSnBase package. A similar filter also applied: (1) unidentified spectrum with msgf e-value < `e evalue_treshold` and (2) spectrum with multiple identification were removed
```{r}
idSummary(msexp.id)
dim(fData(msexp.id)) # before filter
head(fData(msexp.id)[order(fData(msexp.id)[,"scan number(s)"]),])[1:3,cols]
dim(fData(msexp.filter2)) # After Filter
head(fData(msexp.id)[order(fData(msexp.filter2)[,"scan number(s)"]),])[1:3,cols]
```

## Comparing the results
Here is the results from the P3 container. Quantification used "MAX".
```{r}
head(qnt.id[order(qnt.id$accession),])[1:3,cols.p3]
```

To put in contrast, the data aboved is merged with the data from PNNL. `iTRAQ4.114 - iTRAQ4.114` indicates results from P3, while `Ion_114 - Ion_117` indicate results from PNNL.

```{r}
head(qnt.merged[,c(cols.p3, "Ion_114", "Ion_115", "Ion_116", "Ion_117")])[1:3,]
```

To see if the results coherence, correlation among each ions is computed:
```{r}
cor(qnt.merged$iTRAQ4.114, qnt.merged$Ion_114, method = "spearman", "pairwise.complete.obs")
cor(qnt.merged$iTRAQ4.115, qnt.merged$Ion_115, method = "spearman", "pairwise.complete.obs")
cor(qnt.merged$iTRAQ4.116, qnt.merged$Ion_116, method = "spearman", "pairwise.complete.obs")
cor(qnt.merged$iTRAQ4.117, qnt.merged$Ion_117, method = "spearman", "pairwise.complete.obs")
```