### 2016-05-19 Benchmark/Benchmark/itraq-benchmark.Rmd
benchmark

---
title: | 
  | Portable Proteomics Pipeline (P3) Benchmark
author: "Daniel Kristiyanto (daniel.kristiyanto@pnnl.gov)"
date: \today
fontsize: 11pt
output: pdf_document
# bibliography: ../../media/ref.bib
nocite: |
  @domon2006mass, @deutsch2010mass, @MSnba83
csl: ../../media/biomed-central.csl
header-includes:
  \usepackage{fancyhdr}
  \usepackage{graphicx}
  \usepackage{hyperref}
  \pagestyle{fancy}
  \setlength\headheight{28pt}
  \rhead{\includegraphics[width = .2\textwidth]{../../media/pnnllogo.png}}
  \fancyfoot[LE,RO]{kristiyanto/p3:benchmark}
---
\rule{\textwidth}{1pt}

```{r, warning=FALSE, echo=FALSE, include=FALSE, cache=T}
setwd("~/Dropbox/_SPACE/Benchmark/")
file_mzid        <- list.files(path = ".", pattern ="mzid", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)
file_mzml        <- list.files(path = ".", pattern ="mzML$|mzXML$", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)
file_fasta        <- list.files(path = ".", pattern ="fasta", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)

library(MSnbase)
library(mzID)
suppressMessages(library(BiocParallel))

mzid_pnnl <- read.table(file="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_msgfdb_fht.txt", sep="\t", header=T)
mzids_no_switch   <- mzID("TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.mzid")
mzid_p3           <- mzID("TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzid")
msexp.raw         <- readMSData(file_mzml)
msexp_p3          <- addIdentificationData(msexp.raw, id = mzid_p3,verbose=F)
msexp_no_switch   <- addIdentificationData(msexp.raw, id = mzids_no_switch,verbose=F)

p3_table <- fData(msexp_p3)
ns_table <- fData(msexp_no_switch)
```

1. Files:

```{r, echo=F}
show(file_mzml)
show(file_fasta)
```

2. Identification:

`MSGFPlus.jar` was downloaded from \url{https://omics.pnl.gov/software/ms-gf}. Identification was run multiple times, using different computers (Linux, MacOS, within or without Docker), and resulted the same results.

Identification with different switches resulted different results.

```
# Command Line (Without Switch)
java -Xmx8000M -jar P3/MSGFPlus.jar -s \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzML -o \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.mzid \
-d ID_003632_9011437E.fasta

# Command Line (2)
java -Xmx8000M -jar P3/MSGFPlus.jar -s \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzML -o \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzid \
-d ID_003632_9011437E.fasta  \
-t 10ppm -m 0 -inst 1 -e 1 -ti -1,2 -ntt 2 -tda 1 -minLength 6 \
-maxLength 50 -minCharge 2 -maxCharge 5 -n 1 -thread 7 \
-mod MSGFDB_Mods.txt -minNumPeaks 5 -addFeatures 1
```
```{r, echo=F}
# Identification Result
show(file_mzid)
```

3. Result from PNNL
```{r, echo=F}
head(mzid_pnnl)[1:2,c("ResultID", "Scan", "Peptide", "Protein", "MSGFDB_SpecEValue")]
```
4. Result from MSGF+ without any switches (sorted by SpecEValue)

```{r, echo=F}
tmp           <- ns_table
tmp           <- tmp[order(tmp$'ms-gf:specevalue'),]
show(head(tmp[,c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "accession", "pepseq")])[1:3,])

tmp[tmp$`scan number(s)`=="11612" &! is.na(tmp$`scan number(s)`),c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "accession", "pepseq")]
```

Reading the MZID file manually
```
java -Xmx2000M -XX:+UseConcMarkSweepGC -cp ../P3/MSGFPlus.jar edu.ucsd.msjava.ui.MzIDToTsv -i TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.mzid -showQValue 1 -showDecoy 1 -unroll 1
```
```{r, warning=FALSE, echo=FALSE, include=FALSE, cache=T}
setwd("~/Dropbox/_SPACE/Benchmark/")
tsv_ns <- read.table(file ="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.tsv", sep="\t",header = T)

tsv <- read.table(file ="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.tsv", sep="\t",header = T)

```
```{r, echo=F}
head(tsv_ns)[1:3,c("ScanNum", "Peptide", "Protein", "SpecEValue")]
```

5. Result from with switches (2)
```{r, echo=F}
tmp           <- p3_table
tmp           <- tmp[order(tmp$'ms-gf:specevalue'),]
show(head(tmp[,c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "accession", "pepseq")])[1:3,])
tmp[tmp$`scan number(s)`=="11612" &! is.na(tmp$`scan number(s)`),c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "accession", "pepseq")]

```
Reading the MZID file manually
```{r, echo=}
head(tsv)[1:3,c("ScanNum", "Peptide", "Protein", "SpecEValue")]
```