### 2016-06-02 Benchmark/MSGF-Benchmark/msgf-benchmark2.Rmd
benchmark

---
title: | 
  | Portable Proteomics Pipeline (P3) 
  | MSGF Benchmark
author: "Daniel Kristiyanto (daniel.kristiyanto@pnnl.gov)"
date: \today
fontsize: 10pt
output: pdf_document
# bibliography: ../../media/ref.bib
nocite: |
  @domon2006mass, @deutsch2010mass, @MSnba83
csl: ../../media/biomed-central.csl
header-includes:
  \usepackage{fancyhdr}
  \usepackage{graphicx}
  \usepackage{hyperref}
  \pagestyle{fancy}
  \setlength\headheight{28pt}
  \rhead{\includegraphics[width = .2\textwidth]{../../media/pnnllogo.png}}
  \fancyfoot[LE,RO]{kristiyanto/p3:benchmark}
---
\rule{\textwidth}{1pt}
```{r, warning=FALSE, echo=FALSE, include=FALSE, cache=F}
library(MSnbase)
library(mzID)
library(stringr)
```

```{r, warning=FALSE, echo=FALSE, include=FALSE, cache=F}
setwd("~/Dropbox/_SPACE/Benchmark/11")
file_mzid        <- list.files(path = ".", pattern ="mzid", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)
file_mzml        <- list.files(path = ".", pattern ="mzML$|mzXML$", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)
file_fasta        <- list.files(path = ".", pattern ="fasta", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)

mzid_pnnl         <- read.table(file="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_groundtruth_fht.txt", sep="\t", header=T)
frm_docker        <- read.table(file="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.txt", sep="\t", header = T)

pnnl <- mzid_pnnl[, c("Scan", "MSGFDB_SpecEValue", "Peptide")]
pnnl <- pnnl[!duplicated(pnnl$Scan),]
pnnl$Peptide <- substr(as.character(pnnl$Peptide),1,nchar(as.character(pnnl$Peptide))-2)
pnnl$Peptide <- substr(as.character(pnnl$Peptide),3,nchar(as.character(pnnl$Peptide)))
docker       <- frm_docker[,c("Scan","Spec_Evalue","PepSeq")]
```

# 1. Files:

```{r, echo=F}
print("From PNNL: TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_msgfdb_fht.txt")
print("From Docker: TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.txt")
```

# 2. Identification:
Identification used the following switches:
```
# Command Line (2)
java -Xmx8000M -jar P3/MSGFPlus.jar -s \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzML -o \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzid \
-d ID_003632_9011437E.fasta  \
-t 10ppm -m 0 -inst 1 -e 1 -ti -1,2 -ntt 2 -tda 1 -minLength 6 \
-maxLength 50 -minCharge 2 -maxCharge 5 -n 1 -thread 7 \
-mod MSGFDB_Mods.txt -minNumPeaks 5 -addFeatures 1
```

```{r, echo = F, include=F}
evalue = 1e-10
```

#3. Differences. Spec eValue Threshold= `r evalue` 
### a. From PNNL
```{r, echo=F}
c_pnnl = pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print(paste("Remaining peptides:",nrow(c_pnnl)))
# print("====== head ======")
# head(c_pnnl,2)
# print("====== tail ======")
# tail(c_pnnl,2)

```

### b. From Docker
```{r, echo=F}
c_docker = docker[docker$Spec_Evalue<evalue,]
c_docker = c_docker[!is.na(c_docker$Scan),]
c_docker = c_docker[order(c_docker$Spec_Evalue, decreasing = F),]
print(paste("Remaining peptides:",nrow(c_docker)))
# print("====== head ======")
# head(c_docker,2)
# print("====== tail ======")
# tail(c_docker,2)
```

### c. Differences
```{r echo=F}
m_table   <- merge(x=c_pnnl,y=docker,by="Scan", all.x=T)
same      <- m_table[(m_table$MSGFDB_SpecEValue==m_table$Spec_Evalue) & (m_table$Peptide==m_table$PepSeq),]
same      <- same[!is.na(same$Scan),]
print(paste("Exact match:", nrow(same)))
print(paste("From PNNL:", names(same)[2],names(same)[3],". From Docker:", names(same)[4],names(same)[5]))
print(head(same,2))
diff_spec <- m_table[(m_table$MSGFDB_SpecEValue!=m_table$Spec_Evalue) & (m_table$Peptide==m_table$PepSeq),]

diff_spec <- diff_spec[!is.na(diff_spec$Scan),]
print(paste("Different Spec eValue:", nrow(diff_spec)))
tail(diff_spec,3)
diff_pep  <- m_table[(m_table$Peptide!=m_table$PepSeq),]

diff_pep  <- diff_pep[!is.na(diff_pep$Scan),]
print(paste("Different ID and Spec eValue:", nrow(diff_pep)))
tail(diff_pep,3)

unmatched <- m_table[is.na(m_table$Spec_Evalue),]
print(paste("There are",nrow(unmatched),"of unmatched peptides"))
head(unmatched)

```

```{r, echo = F, include=F}
evalue = 1e-15
```

#4. Differences. Spec eValue Threshold = `r evalue` 
### a. From PNNL
```{r, echo=F}
c_pnnl = pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print(paste("Remaining peptides:",nrow(c_pnnl)))
# print("====== head ======")
# head(c_pnnl,2)
# print("====== tail ======")
# tail(c_pnnl,2)

```

### b. From Docker
```{r, echo=F}
c_docker = docker[docker$Spec_Evalue<evalue,]
c_docker = c_docker[!is.na(c_docker$Scan),]
c_docker = c_docker[order(c_docker$Spec_Evalue, decreasing = F),]
print(paste("Remaining peptides:",nrow(c_docker)))
# print("====== head ======")
# head(c_docker,2)
# print("====== tail ======")
# tail(c_docker,2)
```

### c. Differences
```{r echo=F}
m_table   <- merge(x=c_pnnl,y=docker,by="Scan", all.x=T)
same      <- m_table[(m_table$MSGFDB_SpecEValue==m_table$Spec_Evalue) & (m_table$Peptide==m_table$PepSeq),]
same      <- same[!is.na(same$Scan),]
print(paste("Exact match:", nrow(same)))
print(paste("From PNNL:", names(same)[2],names(same)[3],". From Docker:", names(same)[4],names(same)[5]))
print(head(same,2))
diff_spec <- m_table[(m_table$MSGFDB_SpecEValue!=m_table$Spec_Evalue) & (m_table$Peptide==m_table$PepSeq),]

diff_spec <- diff_spec[!is.na(diff_spec$Scan),]
print(paste("Different Spec eValue:", nrow(diff_spec)))
tail(diff_spec,3)
diff_pep  <- m_table[(m_table$Peptide!=m_table$PepSeq),]

diff_pep  <- diff_pep[!is.na(diff_pep$Scan),]
print(paste("Different ID and Spec eValue:", nrow(diff_pep)))
tail(diff_pep,3)

unmatched <- m_table[is.na(m_table$Spec_Evalue),]
print(paste("There are",nrow(unmatched),"of unmatched peptides"))
head(unmatched)
```

```{r, echo = F, include=F}
evalue = 1e-20
```

#5. Differences. Spec eValue Threshold = `r evalue` 
### a. From PNNL
```{r, echo=F}
c_pnnl = pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print(paste("Remaining peptides:",nrow(c_pnnl)))
# print("====== head ======")
# head(c_pnnl,2)
# print("====== tail ======")
# tail(c_pnnl,2)

```

### b. From Docker
```{r, echo=F}
c_docker = docker[docker$Spec_Evalue<evalue,]
c_docker = c_docker[!is.na(c_docker$Scan),]
c_docker = c_docker[order(c_docker$Spec_Evalue, decreasing = F),]
print(paste("Remaining peptides:",nrow(c_docker)))
# print("====== head ======")
# head(c_docker,2)
# print("====== tail ======")
# tail(c_docker,2)
```

### c. Differences
```{r echo=F}
m_table   <- merge(x=c_pnnl,y=docker,by="Scan", all.x=T)
same      <- m_table[(m_table$MSGFDB_SpecEValue==m_table$Spec_Evalue) & (m_table$Peptide==m_table$PepSeq),]
same      <- same[!is.na(same$Scan),]
print(paste("Exact match:", nrow(same)))
print(paste("From PNNL:", names(same)[2],names(same)[3],". From Docker:", names(same)[4],names(same)[5]))
print(head(same,2))
diff_spec <- m_table[(m_table$MSGFDB_SpecEValue!=m_table$Spec_Evalue) & (m_table$Peptide==m_table$PepSeq),]

diff_spec <- diff_spec[!is.na(diff_spec$Scan),]
print(paste("Different Spec eValue:", nrow(diff_spec)))
tail(diff_spec,3)
diff_pep  <- m_table[(m_table$Peptide!=m_table$PepSeq),]

diff_pep  <- diff_pep[!is.na(diff_pep$Scan),]
print(paste("Different ID and Spec eValue:", nrow(diff_pep)))
tail(diff_pep,3)

unmatched <- m_table[is.na(m_table$Spec_Evalue),]
print(paste("There are",nrow(unmatched),"of unmatched peptides"))
head(unmatched)

```

#6. Differences between results from PNNL and DTA file from PNNL.

```{r echo=F, cache=T, warning=F, include=F}
setwd("~/Dropbox/_SPACE/Benchmark/11")
dta_mzid  <- mzID(file="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_dta.mzid")
#dta_mzml  <- readMSData(files = file_mzml[2])
#dta       <- read.table("TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_dta.tsv", sep="\t", header = T)
dta_flt    <- flatten(dta_mzid)
colnames(dta_flt)[2] <- "Scan"

```
MSGF+ was run against `TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_dta.txt` using the same command (10 ppm, without Docker).

```{r echo=F}
dta <- dta_flt[,c("Scan","ms-gf:specevalue","pepseq")]
print(paste("Peptides:",nrow(dta),"with",sum(duplicated(dta$Scan)), "duplicated scan numbers."))
names(dta) <- c("Scan","MSGFDB_SpecEValue","Peptide")
dta <- dta[!duplicated(dta$Scan),]
print(paste("After removing duplicates",nrow(dta), "peptides remained."))

pnnl <- mzid_pnnl[, c("Scan", "MSGFDB_SpecEValue", "Peptide")]
pnnl <- pnnl[!duplicated(pnnl$Scan),]
pnnl$Peptide <- substr(as.character(pnnl$Peptide),1,nchar(as.character(pnnl$Peptide))-2)
pnnl$Peptide <- substr(as.character(pnnl$Peptide),3,nchar(as.character(pnnl$Peptide)))

m_table <- merge(x=pnnl, y=dta, by=c("Scan"), all=T)

same <- m_table[(m_table$MSGFDB_SpecEValue.x ==m_table$MSGFDB_SpecEValue.y) & (m_table$Peptide.x == m_table$Peptide.y),]

diff_spec <- m_table[(m_table$MSGFDB_SpecEValue.x !=m_table$MSGFDB_SpecEValue.y) & (m_table$Peptide.x == m_table$Peptide.y),]

diff_pep <- m_table[(m_table$Peptide.x != m_table$Peptide.y),]
```